Static Discharge and Demotion of Low-Level, Higher-Order Contracts

### 2013-11-07 pldi2014/paper.tex
First three pages are in final draft form.

\title{Static Discharge and Demotion of Higher-Order Contracts}
%\subtitle{Subtitle Text, if any}

@@

@@

Consider, for instance, the program
\begin{schemedisplay}
(define/contract (fac n a)
  (nonnegative-integer/c positive-integer/c . -> . positive-integer/c)
  (if (= n 0)
      a
      (fac (- n 1) (* n a))))
\end{schemedisplay}
which attaches dynamic domain and range contracts to the tail-recursive factorial.
This contracted procedure behaves as if it were written
\begin{schemedisplay}
(define (fac n a)
  (positive-integer/c
   ((LAMBDA (n a)
      (if (= n 0)
          a
          (fac (- n 1) (* n a))))
    (nonnegative-integer/c n)
    (positive-integer/c a))))
\end{schemedisplay}
where \scheme{positive-integer/c} and \scheme{nonnegative-integer/c} are procedures which raise an error if their argument does not belong to the domain reflected by the procedure's name.
Written in this way, it is apparent that \scheme{fac} is not tail-recursive.
Instead, the context grows with every recursive call so that \scheme{(fac 6 1)} reduces as
\begin{schemedisplay}
(fac 6 1)
(positive-integer/c (fac 5 6))
(positive-integer/c (positive-integer? (fac 4 30)))
...
(positive-integer/c ... (positive-integer/c (fac 0 720)))
\end{schemedisplay}
whence the result will be validated six times---and \scheme{n} times for \scheme{(fac n 1)}.
The attachment of these contracts has transformed a constant-space algorithm into a linear-space one and, while the time complexity hasn't suffered asymptotically, the accumulated checks represent a real (and, in this case, unnecessary) cost.
Thus, a simple contract ascription can upset our performance expectations for even simple procedures.

As contract specifications become more elaborate, they become more susceptible to error.
In contract systems, such as Racket's~\cite{plt-tr1}, which allow arbitrary predicate procedures written in Racket itself to serve as specifications, this error is effected by the programmer.
It is only natural, then, to contract these procedures as well, an act which aides error diagnosis but compounds the run-time cost of contracts.
In these systems, execution time of the program proper can be dwarfed by that of the enforcement of its specification (in one case accounting for 96\% of total execution time) \cite{strickland2012chaperones}.

The contract cost problem has been studied extensively.
Herman et al.~\cite{herman2010space} observed the potential accumulation of casts in the blame calculus---similar in principle to the accumulation of guards in our contracted \scheme{fac}---and applied the coercion calculus of Henglein~\cite{henglein1994dynamic} to compact them.
Siek and Wadler~\cite{siek2010threesomes} consider the same problem and apply \emph{threesomes}.
Siek et al.~\cite{siek2006gradual} explicate type systems which allow \emph{gradual typing} in which they provide a low-cost path from dynamically- to statically-checked invariants.
Flanagan and Knowles~\cite{flanagan2006hybrid} introduce \emph{hybrid type checking} in which static analysis is used to establish contract compliance statically where possible and insert dynamic checks where not.
Using this approach, a disciplined programmer enjoys the early safety guarantees and performance characteristics of statically-verified code or suffers the cost of belated error detection of dynamically-checked specifications.

Modular development can impede these analyses because the greatest amount of knowledge a module can have about its imports \emph{is} their behavioral contract.\footnote{Indeed, it can be less than this, as when the behavioral specification is not codified entirely within the program.}
For example, recognizing that contracts constitute total formal knowledge, Tobin-Hochstadt and Van Horn \cite{tobin2012higher} conduct an intra-module analysis which identifies extra-module references with opaque values that satisfy their contract.
% also, contracts are attached to values that flow across module boundaries.
Nevertheless, static analyses typically perform best when given whole programs.
For instance, the whole-program ML compiler MLton \cite{weeks2006whole} is able to perform thorough inter-procedural and inter-modular analysis and effect significant program optimizations.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Might and Prabhu~\cite{might2009interprocedural} allowing us to judge the safe removal of dynamic contracts \emph{even when the creation of other contracts depends on them}.
The expressiveness of Racket's contract language admits dynamic checks which diverge, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-created contracts.

, which disregards some latent structure provided by the analysis.
Specifically, the pushdown system characterization of an abstract interpretation immediately admits a finite description of an unbounded stack as a regular expression.
We use this structure to approximate blame on a misbehaving program.
Using , coupled with a termination analysis based on Sereni's~\cite{sereni2007termination}, we devise a blame analysis which allows us to soundly and statically discharge contracts.

We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
We conclude by discussing the future application and direction of this work in Section 8.

@@

@@

The contracted tail-recursive factorial procedure expressed as a chaperone is
\begin{schemedisplay}
(define fac (chaperone-procedure
             (LAMBDA (n a)
               (if (= n 0)
                   a
                   (fac (- n 1) (* n a))))
             (LAMBDA (n a)
               (if (and (nonnegative-integer? a)
                        (positive-integer? a))
                   (values (LAMBDA (a)
                             (if (positive-integer? a)
                                 a
                                 (raise positive blame ...)))
                           n a)
                   (raise negative blame ...)))))
\end{schemedisplay}

@@

\newcommand{\stxclass}[4]{$#1\in\mathbf{#2}$ &=& #3 & #4}
\newcommand{\stxclasscont}[2]{&$|$& #1 & #2}
%\newcommand{\fstxclass}[3]{$#1$ &=& #2

@@

\begin{tabular}{ r | r | l | l }
\stxclass{c}{Call}{\scheme|(ae ae ...)|$^\ell$}{application}\\
\stxclass{e}{Exp}{\scheme|(let ((x ...) c) e)|}{let}\\
\stxclasscont{\scheme|(if ae e e)|}{if}\\
\stxclasscont{\scheme|ae|}{}\\
\stxclass{ae}{AExp}{$\mathit{lam}$ $|$ x $|$ p $|$ i $|$ b}{atomic expressions}
%\stxclass{\mathit{lam}}{Lam}{\scheme|(LAMBDA (x ...) e)|}{lambda terms}
%\stxclass{p}{Prim}{\scheme|impersonate-operator| $|$ \scheme|<| $|$ \scheme|>| $|$ %\scheme|=| $|$ \scheme|+| $|$ \scheme|-|}{}
\end{tabular}

\begin{verbatim}
 c\in Call &= (ae ae ...)^l               &\text{application}\\
 e\in Exp  &= (let ((x ...) c) e)         &\text{let}\\
           &| (if ae e e)                 if\\
           &| ae
ae\in AExp &=  lam | x | p | i | b
lam\in Lam &= (lambda (x ...) e)
 p &= impersonate-operator | < | > | = | + | -
   &| chaperone-operator | not | values | raise
   &| integer? | boolean? | operator?
 x &= variables
 i &= 0 | 1 | -1 | 2 | -2 | ...
 b &= \#t | \#f
\end{verbatim}

%\begin{align*}
%c &= \appe{\ae}{\ae\,\dots} && \text{application}\\
%e &= \lete{x\,\dots}{c}{e} && \text{let}\\
%  &|  \letrece{x\,\dots}{c}{e} && \text{letrec}\\
%  &|  \ife{\ae}{e}{e} && \text{if}\\
%  &   \chae{\ae}{e} && \text{chaperone-operator}\\
%  &   \impe{\ae}{e} && \text{impersonate-operator}\\
%\ae &= \lame{x\,\dots}{e}\,|\,x\,|\,p\,|\,i\,|\,b && tsh\\
%p &= values | = | < | > | + | - | not | integer? | boolean? | operator? | raise && \\
%x &= ha & \\
%i &= 0|1|-1|2|-2|\dots\\
%b &= \true\,|\,\false\\
%\end{align*}

%\begin{tabular}{rrl}
%$c$ & = & $\appe{\ae}{\ae\,\dots}$\\
%$e$ & = & $\lete{x\,\dots}{c}{e}$\\
%g & h & i
%\end{tabular}
\caption{The syntax of \chapcalc.}
\end{figure}

@@

% raise integer? boolean? not operator? + - * = < > values
% need primitives to allow contracts to deal higher orderly
% ae := (lambda s e) x primitive integer boolean

% (ae ae ...)
% (app-values ae ae)
% (chaperone-operator ae ae ae)
% (impersonate-operator ae ae ae)
% (let ([s call]) e)
% (letrec ([s call]) e) IS IT POSSIBLE TO LIFT AN EXPRESSION OUT THAT SHOULDN'T BE?
% (if ae e e)
% (lambda s e)

@@

\newcommand{\ks}[0]{\kappa s}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\Sigma_e(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\Sigma_v(#1,#2,#3)}
\newcommand{\sapp}[4]{\Sigma_a(#1,#2,\ell,#3,#4)}
\newcommand{\scha}[6]{\Sigma_d(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdimpz}[6]{\Sigma_{imp_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdimpo}[5]{\Sigma_{imp_1}(#1,#2,#3,#4,#5)}
\newcommand{\sdchaz}[6]{\Sigma_{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdchao}[5]{\Sigma_{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[3]{\Sigma_!(#1,#2,#3)}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\red}[2]{\begin{align*}& #1\\\rr\, & #2\end{align*}}

@@

A recursive definition of the form
\begin{schemedisplay}
(letrec ([(x_1 x_2 ... x_n) (values e_1 e_2 ... e_n)])
  e)
\end{schemedisplay}
can be written
\begin{schemedisplay}
(let ([(x_1 x_2 ... x_n) (Y_n (LAMBDA (x_1 x_2 ... x_n) e_1)
                              (LAMBDA (x_1 x_2 ... x_n) e_2)
                              ...
                              (LAMBDA (x_1 x_2 ... x_n) e_n))])
  e)
\end{schemedisplay}
where \scheme{Y_n} is defined as
\begin{schemedisplay}
(LAMBDA (f_1 f_2 ... f_n)
  ((LAMBDA (fp_1 fp_2 ... fp_n)
     (values (fp_1 fp_1 fp_2 ... fp_n)
             (fp_2 fp_1 fp_2 ... fp_n)
             ...
             (fp_n fp_1 fp_2 ... fp_n)))
   (LAMBDA (fp_1 fp_2 ... fp_n)
     (LAMBDA (x) ((f_1 (fp_1 fp_1 fp_2 ... fp_n)) x)))
   (LAMBDA (fp_1 fp_2 ... fp_n)
     (LAMBDA (x) ((f_2 (fp_2 fp_1 fp_2 ... fp_n)) x)))
   ...
   (LAMBDA (fp_1 fp_2 ... fp_n)
     (LAMBDA (x) ((f_n (fp_n fp_1 fp_2 ... fp_n)) x)))))
\end{schemedisplay}

@@

expression: forms + primitives + booleans + numbers + labels + variables
value: primitive + boolean + number + closure + impersonator + chaperone
semantics: stack frame + stack (sequence of frames) + store + address + environment as finite map
$\Gamma^{*}$

@@

In difference, our machine is defined in terms of multiple state variants.
\begin{itemize}
\item An \emph{evaluation} state $\Sigma_{e}(\langle\gamma,\dots\rangle,\sigma,\rho,e)$ corresponds to a traditional CESK state.
(The order of the components is reversed, however, to aid factoring out common components from among the variants.)

@@

\item An \emph{apply} state $\Sigma_{a}(\langle\gamma,\dots\rangle,\sigma,f,\mathbf{v})$ represents a point of application and dispatches on the operator type (closure, impersonator, or primitive).
\item A \emph{value} state $\Sigma_{v}(\langle\gamma,\dots\rangle,\sigma,\mathbf{v})$ represents a point of return and dispatches on the type of the top continuation frame.
If the continuation is empty, the constituent values are the results of the program.
\item An \emph{error} state $\Sigma_!(\langle\gamma,\dots\rangle,\sigma,b)$, represents the arisal of an error and, as the result of the program, may include blame information.
\item Finally, we define a set of \emph{dummy} states which splits a transition which would otherwise pop and push a frame simulaneously into a sequence of two transitions which performs each separately.
(This eases further the transition to a pushdown system.)
\end{itemize}

@@

Because higher-order contracts often refine domains, the choice of abstraction function can have significant ramifications for the precision of the analysis.
For instance, an analysis of the program
\begin{schemedisplay}
(letrec ([(fac) (values (LAMBDA (n)
                          (let ([test0 (= n 0)])
                            (if test0
                                1
                                (let ([(n-1) (- n 1)])
                                  (let ([(rec) (fac n-1)])
                                    (* n rec)))))))])
  (let ([(fac) (-> fac nonnegative? positive?)])
    (fac 5)))
\end{schemedisplay}
using a singleton domain [?] for integers will fail to discharge the chaperone.
A domain which distinguishes values coincident with the edges of the contract would be able to.

>>>>

\title{Static Discharge and Demotion of Low-Level, Higher-Order Contracts}
%\subtitle{Subtitle Text, if any}

@@

\setspecialsymbol{->}{$\rightarrow$}

@@

Consider, for instance, the program
\begin{schemedisplay}
(define/contract (fact n a)
  (nonnegative-integer/c positive-integer/c . -> . positive-integer/c)
  (if (= n 0)
      a
      (fact (- n 1) (* n a))))
\end{schemedisplay}
which attaches dynamic domain and range contracts to the tail-recursive factorial.
This contracted procedure behaves as if it were written
\begin{schemedisplay}
(define (fact n a)
  (positive-integer/c
   ((LAMBDA (n a)
      (if (= n 0)
          a
          (fact (- n 1) (* n a))))
    (nonnegative-integer/c n)
    (positive-integer/c a))))
\end{schemedisplay}
where \scheme{positive-integer/c} and \scheme{nonnegative-integer/c} are procedures which raise an error if their argument does not belong to the domain reflected by the procedure's name.
Written in this way, it is apparent that \scheme{fac} is not tail-recursive.
Instead, the context grows with every recursive call so that \scheme{(fact 6 1)} reduces as
\begin{schemedisplay}
(fact 6 1)
(positive-integer/c (fact 5 6))
(positive-integer/c (positive-integer/c (fact 4 30)))
...
(positive-integer/c ... (positive-integer/c (fact 0 720)))
(positive-integer/c ... (positive-integer/c 720))
\end{schemedisplay}
whence the result will be validated six times---and \scheme{n} times for \scheme{(fact n 1)}.
The attachment of these contracts has transformed a constant-space algorithm into a linear-space one and, while the time complexity hasn't suffered asymptotically, the accumulated checks represent a real (and, in this case, unnecessary) cost.
Thus, a simple contract ascription can upset our performance expectations for even simple procedures.

As contract specifications become more elaborate, they become more susceptible to error.
In contract systems, such as Racket's~\cite{plt-tr1}, which allow arbitrary predicate procedures written in the source language itself to serve as specifications, this error is effected by the programmer.
It is only natural, then, to contract these procedures as well, an act which aides error diagnosis but compounds the run-time cost of contracts.
In these systems, execution time of the program proper can be dwarfed by that of the enforcement of its specification (in one case accounting for 96\% of total execution time) \cite{strickland2012chaperones}.

The moderation of these costs has been studied extensively.
Herman et al.~\cite{herman2010space} observed the potential accumulation of casts in the blame calculus---similar in principle to the accumulation of guards by our contracted \scheme{fact} procedure---and applied the coercion calculus of Henglein~\cite{henglein1994dynamic} to compact them.
Siek and Wadler~\cite{siek2010threesomes} consider the same problem and apply their own solution of \emph{threesomes}.
Siek et al.~\cite{siek2006gradual} explicate type systems which allow \emph{gradual typing} in which they provide a low-cost path from dynamically- to statically-checked invariants.
Flanagan and Knowles~\cite{flanagan2006hybrid} introduce \emph{hybrid type checking} in which static analysis is used to establish contract compliance statically where possible and insert dynamic checks where not.
%Using this final approach, a disciplined programmer enjoys the early safety guarantees and performance characteristics of statically-verified code or suffers the cost of belated error detection of dynamically-checked specifications.

Unfortunately, modular development can impede these analyses because the greatest amount of knowledge a module can have about its imports \emph{is} their behavioral contract.\footnote{Indeed, it can be less than this, as when the behavioral specification is not codified entirely within the program.}
For example, recognizing that contracts constitute total formal knowledge, Tobin-Hochstadt and Van Horn \cite{tobin2012higher} conduct an intra-module analysis which identifies extra-module references with opaque values that satisfy their contract.
% also, contracts are attached to values that flow across module boundaries.
Nevertheless, static analyses typically perform best when given whole programs.
For instance, the whole-program ML compiler MLton \cite{weeks2006whole} is able to perform thorough inter-procedural and inter-modular analysis and effect significant program optimizations.
We similarly expect the establishment of certain properties to become more accessible as code knowledge increases.
%We expect the emergence of the possibility of more properties being proved about the code as more knowledge about it becomes available.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Might and Prabhu~\cite{might2009interprocedural} allowing us to judge the safe removal of dynamic contracts \emph{even when the installation of other contracts depends on them}.
The expressiveness of Racket's contract language admits dynamic checks which diverge, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-installed contracts.

We proceed as follows:
We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
We conclude by discussing the future application and direction of this work in Section 8.

@@

\subsection{Mapping from contracts to impersonators}

@@

The contracted tail-recursive factorial procedure expressed as a chaperone is
\begin{schemedisplay}
(define fact (chaperone-procedure
             (LAMBDA (n a)
               (if (= n 0)
                   a
                   (fact (- n 1) (* n a))))
             (LAMBDA (n a)
               (if (and (nonnegative-integer? a)
                        (positive-integer? a))
                   (values (LAMBDA (a)
                             (if (positive-integer? a)
                                 a
                                 (raise positive blame ...)))
                           n a)
                   (raise negative blame ...)))))
\end{schemedisplay}

@@

\newcommand{\stxclass}[4]{$#1\in\mathbf{#2}$ &::=& #3 & #4}
\newcommand{\stxclasscont}[2]{&$|$& #1 & #2}

@@

\begin{tabular}{ r r l l }
\stxclass{c}{Call}{\scheme|(ae ae ...)|$^\ell$}{application}\\
\stxclass{e}{Exp}{\scheme|(let ((x ...) c) e)|}{let}\\
\stxclasscont{\scheme|(if ae e e)|}{if}\\
\stxclasscont{\scheme|ae|}{}\\
\stxclass{\mathit{ae}}{AExp}{$\mathit{lam}$ $|$ $p$ $|$ $x$ $|$ $i$ $|$ $b$}{atomic expressions}\\
\stxclass{\mathit{lam}}{Lam}{\scheme|(LAMBDA (x ...) e)|}{lambda terms}\\
\stxclass{p}{Prim}{\scheme|impersonate-operator| $|$ \scheme|<| $|$ \scheme|>| $|$ \scheme|=| $|$ \scheme|+| $|$ \scheme|-|}{primitives}\\
\stxclasscont{\scheme|chaperone-operator| $|$ \scheme|values| $|$ \scheme|raise|}{}\\
\stxclasscont{\scheme|not| $|$ \scheme|operator?| $|$ \scheme|integer?| $|$ \scheme|boolean?|}{}\\
\stxclass{x}{Var}{an infinite set of variables}{}\\
\stxclass{i}{Int}{$0\,|\,1\,|\,-1\,|\,2\,|\,-2\,|\,\dots$}{integers}\\
\stxclass{b}{Bool}{$\#t\,|\,\#f$}{booleans}
\end{tabular}

\caption{The syntax of \chapcalc.}
\end{figure}

@@

@@

\newcommand{\ks}[0]{\langle\gamma_1,\dots,\gamma_n\rangle}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\mathbf{eval}(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\mathbf{value}(#1,#2,#3)}
\newcommand{\sapp}[4]{\mathbf{apply}(#1,#2,\ell,#3,#4)}
\newcommand{\scha}[6]{\Sigma_d(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdimpz}[6]{\mathbf{imp_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdimpo}[5]{\mathbf{imp_1}(#1,#2,#3,#4,#5)}
\newcommand{\sdchaz}[6]{\mathbf{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\sdchao}[5]{\mathbf{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[3]{\mathbf{error}(#1,#2,#3)}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\red}[2]{\begin{align*}& #1\\\rr\, & #2\end{align*}}

@@

%A recursive definition of the form
%\begin{schemedisplay}
%(letrec ([(x_1 x_2 ... x_n) (values e_1 e_2 ... e_n)])
%  e)
%\end{schemedisplay}
%can be written
%\begin{schemedisplay}
%(let ([(x_1 x_2 ... x_n) (Y_n (LAMBDA (x_1 x_2 ... x_n) e_1)
%                              (LAMBDA (x_1 x_2 ... x_n) e_2)
%                              ...
%                              (LAMBDA (x_1 x_2 ... x_n) e_n))])
%  e)
%\end{schemedisplay}
%where \scheme{Y_n} is defined as
%\begin{schemedisplay}
%(LAMBDA (f_1 f_2 ... f_n)
%  ((LAMBDA (fp_1 fp_2 ... fp_n)
%     (values (fp_1 fp_1 fp_2 ... fp_n)
%             (fp_2 fp_1 fp_2 ... fp_n)
%             ...
%             (fp_n fp_1 fp_2 ... fp_n)))
%   (LAMBDA (fp_1 fp_2 ... fp_n)
%     (LAMBDA (x) ((f_1 (fp_1 fp_1 fp_2 ... fp_n)) x)))
%   (LAMBDA (fp_1 fp_2 ... fp_n)
%     (LAMBDA (x) ((f_2 (fp_2 fp_1 fp_2 ... fp_n)) x)))
%   ...
%   (LAMBDA (fp_1 fp_2 ... fp_n)
%     (LAMBDA (x) ((f_n (fp_n fp_1 fp_2 ... fp_n)) x)))))
%\end{schemedisplay}

@@

@@

Our machine is defined in terms of multiple state variants.
\begin{itemize}
\item An \emph{evaluation} state $\sexp{\ks}{\sigma}{\rho}{e}$ corresponds to a traditional CESK state.
(The order of the components is reversed, however, to aid factoring out common components from among the variants.)

@@

\item A state $\sapp{\ks}{\sigma}{f}{\mathbf{v}}$ represents a point of application and dispatches on the operator type (closure, impersonator, or primitive).
\item A state $\sval{\ks}{\sigma}{\mathbf{v}}$ represents a point of return and dispatches on the type of the top continuation frame.
If the continuation is empty, the constituent values are the results of the program.
\item A state $\serr{\ks}{\sigma}{b}$, represents the arisal of an error and, as the result of the program, may include blame information.
\item Finally, we define a set of \emph{dummy} states which splits a transition which would otherwise pop and push a frame simulaneously into a sequence of two transitions which performs each separately.
(This eases further the transition to a pushdown system.)
\end{itemize}

%expression: forms + primitives + booleans + numbers + labels + variables
%value: primitive + boolean + number + closure + impersonator + chaperone
%semantics: stack frame + stack (sequence of frames) + store + address + environment as finite map
%$\Gamma^{*}$

%\begin{verbatim}
%\Sigma_e(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},\hat{\rho},e)
%\Sigma_a(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},\ell,\hat{f},\hat{\mathbf{v}})
%\Sigma_v(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},\hat{\mathbf{v}})
%\Sigma_!(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},b)
%\Sigma_0(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},\ell,\hat{f},\hat{w},\hat{\mathbf{v}})
%\Sigma_1(\langle\hat{\gamma},\dots\rangle,\hat{\sigma},\ell,\hat{w},\hat{\mathbf{v}})

%\hat{\sigma}\in \hat{Addr}\mapsto powerset(\hat{V})
%\hat{\rho}\in Var\mapsto\hat{Addr}
%\alpha\in\hat{Addr}
%f,w\in\hat{Op}=\hat{Clo}\cup\hat{Imp}\cup\hat{Prim}
%v\in\hat{V}=\hat{Op}\cup\hat{Bool}\cup\hat{Int}
%\end{verbatim}

@@

Because higher-order contracts often refine domains, the choice of abstraction function can have significant ramifications for the precision of the analysis.
For instance, an analysis of the program
\begin{schemedisplay}
(letrec ([(fact) (values (LAMBDA (n)
                          (let ([test0 (= n 0)])
                            (if test0
                                1
                                (let ([(n-1) (- n 1)])
                                  (let ([(rec) (fact n-1)])
                                    (* n rec)))))))])
  (let ([(fact) (-> fact nonnegative? positive?)])
    (fact 5)))
\end{schemedisplay}
using a singleton domain [?] for integers will fail to discharge the chaperone.
A domain which distinguishes values coincident with the edges of the contract would be able to.
